### 2023-07-19 writing/manuscript.Rmd
set up public repo post rev 2

---
bibliography: themusiclab.bib
csl: pnas.csl
header-includes:
- \usepackage[left]{lineno}
- \usepackage{ragged2e}
- \usepackage{caption}
- \usepackage{longtable}
- \usepackage[labelformat = empty]{caption}
- \usepackage{afterpage}
- \usepackage{mdframed}
- \usepackage{fontenc}
- \usepackage{soul}
- \usepackage{xcolor}
- \usepackage[symbol]{footmisc}
- \definecolor{bleu}{HTML}{2200cc}
- \renewcommand{\thefootnote}{\fnsymbol{footnote}}
notes-after-punctuation: no
urlcolor: bleu
linkcolor: bleu
link-citations: yes
output:
  pdf_document:
    number_sections: yes
  word_document: default
---

@@

```{r libraries}

library(pacman)
p_load(
  here,
  nlme,
  lme4,
  multcomp,
  RColorBrewer,
  lmerTest,
  gt,
  scales,
  ggpubr,
  geodist,
  cowplot,
  patchwork,
  ggsignif,
  broom,
  broom.mixed,
  ggeffects,
  ggtext,
  ggnewscale,
  glue,
  rgeos,
  rnaturalearth,
  rnaturalearthdata,
  sf,
  tidyverse,
  kableExtra
)

@@

```{r load_data}

fulldata <- read_csv(here("results", "FFfull.csv"), 
                     col_types = "fiiffiifcccccddddddddddffffddfcccffff") %>% 
  left_join(.,
            read_csv(here("data", "countries.csv"),
                     col_types = "ccc"), 
            by = "natcountry") %>% 
  mutate(songfunction = factor(songfunction, levels = c("danc", "baby", "heal", "love")))

# data storage lists
mods <- list()

```{r include=FALSE}
# some participants slipped thru the Qualtrics exclusion criteria. The code below identifies 8 participants whose IP address did not match the country they reported being in. For the public repository for this data, we have removed identifying info such as IP addresses. As such, these participants have been manually removed.
# 
# mismatch_ips <- fulldata %>% 
#   select(indx,geo_country_name, natcountry, listener_subregion) %>% distinct() %>% 
#   mutate(oops = ifelse(geo_country_name != natcountry, 1,0)) %>% 
#   filter(oops == 1 & natcountry ! = "Czech Republic" &  natcountry != "Russian Federation") %>% distinct()

mismatch_ips <- c(1994, 2001, 2005, 2010, 2159, 4822, 4999, 5036) # manually remove mismatched IPs

fulldata <- fulldata %>% 
  filter(!indx %in% mismatch_ips)

```{r wrangling}

# generate song level data for Ind. cohort confirmatory analyses
webdat <- fulldata %>% 
  filter(study == "web") %>% 
  group_by(song, songfunction) %>% 
  # averaging scores
  summarize(across(
    c(danc, heal, baby, love, achi, visi),
    ~ mean(.x, na.rm = T),
    .names = "mean_{.col}")) %>% 
  ungroup() %>% 
  # z-scoring
  mutate(across(
    c(mean_danc, mean_baby, mean_heal, mean_love),
    ~ scale(.x),
    .names = "zm_{.col}"
  )) %>% 
  rename_with(~ str_remove(.x, "mean_"), contains("zm")) %>% 
  # dummy variables
  mutate(var = 1, temp = songfunction,
         songfunction = factor(songfunction, levels = c("danc", "heal", "love", "baby"))) %>% 
  pivot_wider(names_from = temp, values_from = var, values_fill = 0)

# generate song level data for SS cohort confirmatory analyses
fieldraw <- fulldata %>% 
  filter(study == "field")

fielddat <- fieldraw %>% 
  mutate(across(
    c(danc, heal, baby, love),
    # the recoding is because the field society responses were originally coded as
    # 11,12,13 instead of 1,2,3, to avoid confusing the scale with the 4-point industrialised cohort scale. 
    ~ recode(.x, `11` = 1, `12` = 2, `13` = 3)
  )) %>% 
  group_by(song, songfunction) %>% 
  # averaging scores
  summarize(across(
    c(danc, heal, baby, love, achi, visi, stor, dead),
    ~ mean(.x, na.rm = T),
    .names = "mean_{.col}")) %>% 
  ungroup() %>% 
  # z-scoring
  mutate(across(
    c(mean_danc, mean_baby, mean_heal, mean_love),
    ~ scale(.x),
    .names = "zm_{.col}"
  )) %>% 
  rename_with(~ str_remove(.x, "mean_"), contains("zm")) %>% 
  # dummy variables
  mutate(var = 1, temp = songfunction) %>% 
  pivot_wider(names_from = temp, values_from = var, values_fill = 0)

webraw <- fulldata %>% 
  filter(study == "web") %>% 
  select(-c("stor", "dead")) %>%
  mutate(
    langshare = ifelse(songlang_fam == natlang_fam, "Shared", "Not shared"),
    geoshare_sub = ifelse(song_subregion == listener_subregion, "Shared", "Not shared"),
    geoshare_reg = ifelse(song_region == listener_region, "Shared", "Not shared")
  ) %>% 
  # z-scoring
  mutate(across(
    c(danc, baby, heal, love),
    ~ scale(.x),
    .names = "zm_{.col}"
  ),
  # adding distance variable
  dist = geodist_vec(song_latitude,
                    song_longitude,
                    geo_latitude,
                    geo_longitude, 
                    paired=T,
                    measure="haversine")) %>% 
  mutate(z_dist = scale(dist),
         log_dist = log1p(dist))

# combined data
combined_data <- left_join(
  webdat %>% select(song, songfunction, 
                    web_danc = mean_danc,
                    web_heal = mean_heal,
                    web_baby = mean_baby,
                    web_love = mean_love), 
  fielddat %>% select(song, songfunction, 
                      field_danc = mean_danc,
                      field_heal = mean_heal,
                      field_baby = mean_baby,
                      field_love = mean_love), 
  by = c("song", "songfunction"))

```{r}
# store demographic info about ind. and SS cohorts

info <- list()
info$web <- map(c("indx", "natlang", "natlang_fam", "natcountry") %>%
                  set_names(c("n","lang","langfam","country")),
                ~ n_distinct(webraw[[.x]]))
info$field$n <- fieldraw$indx %>% n_distinct

```{r}

# Function to run linear models and spit out formatted list
mod_extracter <- function(outcome_var, contrasts, target, data_x) {
  out <- list()
  x <- reformulate(c("danc + baby + love + heal - 1"), response = outcome_var)
  mod <- lm(x, data = data_x)
  out$mod <- mod %>% glance
  out$coef <- mod %>% tidy %>% split(.$term)
  lht <- glht(mod, linfct = contrasts)
  out$lht <- lht %>%
    summary(., adjusted("bonferroni")) %>% tidy %>%
    bind_cols(., lht %>% confint %>% tidy %>% select(conf.low, conf.high)) %>% 
    mutate(contrast = snakecase::to_snake_case(contrast)) %>% split(.$contrast)
  x <- reformulate(c(target), response = outcome_var)
  out$lht$diff <- lm(x, data = data_x) %>% tidy %>%
    mutate(term = snakecase::to_snake_case(term)) %>% split(.$term)

  return(out)
}

mods$danc <- mod_extracter("zm_danc", c("baby - danc = 0", 
                     "heal - danc = 0",
                     "love - danc = 0"), 
                     "danc", webdat)
mods$baby <- mod_extracter("zm_baby", c("danc - baby = 0",
                     "heal - baby = 0",
                     "love - baby = 0"),
                     "baby", webdat)
mods$heal <- mod_extracter("zm_heal", c("baby - heal = 0",
                     "danc - heal = 0",
                     "love - heal = 0"),
                     "heal", webdat)
mods$love <- mod_extracter("zm_love", c("baby - love = 0",
                     "heal - love = 0",
                     "danc - love = 0"),
                     "love", webdat)

\raggedright
\LARGE
\textbf{Mutual intelligibility in musical communication}

@@

\justifying
\normalsize
Lidya Yurdum^1,2,$\ast$^, Manvir Singh^3^, Luke Glowacki^4^, Tom Vardy^5^, Quentin D. Atkinson^5^, Courtney B. Hilton^2,5^, Disa Sauter^1^, Max M. Krasnow^6^, \& Samuel A. Mehr^2,5,$\ast$^

\small
^1^Department of Psychology, University of Amsterdam, Amsterdam 1018WT, Netherlands.  
^2^Haskins Laboratories, Yale University, New Haven, CT 06511, USA.  
^3^Institute for Advanced Study in Toulouse, 31080 Toulouse Cedex 6, France.  
^4^Department of Anthropology, Boston University, Boston, MA 02215, USA.  
^5^School of Psychology, University of Auckland, Auckland 1010, New Zealand.  
^6^Division of Continuing Education, Harvard University, Cambridge, MA 02138, USA.

\*Corresponding authors. E-mail: [lidya.yurdum\@yale.edu](mailto:lidya.yurdum@yale.edu){.email}, [sam\@yale.edu](mailto:sam@yale.edu){.email}

@@

```{=tex}
\normalsize
\begin{mdframed}[backgroundcolor=gray!20]
Despite the variability of music across cultures, some types of human songs share acoustic characteristics with one another. For example, dance songs tend to be loud and rhythmic and lullabies tend to be quiet and melodious. Human perceptual sensitivity to the behavioural contexts of songs on the basis of these acoustic features raises the possibility that basic properties of music are mutually intelligible, independent of linguistic or cultural content. Whether these effects reflect a universal perceptual phenomenon, however, is unclear, because prior studies focus almost exclusively on English-speaking participants, a group that is not representative of humans, writ large. Here we report shared intuitions concerning the behavioural contexts of unfamiliar songs produced in unfamiliar languages, in participants living in Internet-connected industrialised societies ($n = 5,516$ native speakers of 28 languages) or smaller-scale societies with limited access to global media ($n = 116$ native speakers of 3 non-English languages). Participants listened to songs randomly selected from a representative sample of human vocal music, originally used in four behavioural contexts, and rated the degree to which they believed the song was used for each context. Listeners in both industrialised and smaller-scale societies reliably inferred the contexts of dance songs, lullabies, and healing songs, but not love songs. Within and across the cohorts, inferences were mutually consistent. Further, increased linguistic or geographical proximity between listeners and singers only minimally increased the accuracy of the inferences. These results demonstrate that the behavioural contexts of three common forms of music are mutually intelligible across cultures and imply that musical diversity, shaped by cultural evolution, is nonetheless grounded in some universal principles.

@@

Like many other animals, humans use vocalisations to convey their intentions and affective states [@Morton1977; @Pisanski2022]. Such vocalisations would be meaningless in a world where members of one's own — or other — species could not interpret these signals in a useful way. Indeed, many animal and human vocalisations are not arbitrary but instead display systematic relationships between their acoustic form and their behavioural function [@Endler1993; @Fitch2002; @Pisanski2022]. For instance, the human scream is unlikely to have evolved arbitrarily as a means of communicating distress and urgency: rather, a scream involves extreme high frequencies [@Pisanski2020] and acoustic roughness [@Arnal2015] that set it apart from regular verbal communication, and make it explicitly appropriate for the behavioural function of grabbing attention.

@@

Systematic form-function relationships also apply to more complex vocalisations. Song is a human universal characterised by rich variability within and across cultures [@Mehr2019; @Nettl1964; @Lomax1968]. Some of the behavioural contexts in which songs are used, however, are conspicuously similar around the globe, such as singing to soothe fussy infants, or singing to coordinate dancing [@Trehub1993; @Trehub1993a; @Mehr2018a; @Mehr2019; @Hilton2022a; @Hilton2022b; @Singhinpressa]. Songs used for specific functions in specific behavioural contexts tend to display stereotyped acoustic features: for example, dance songs tend to share clearly accented and predictable beat structures. As with other types of vocalisations, form-function patterns in human song may originate from our evolved psychology, perceptual biases, or unique social environment [@Hagen2003; @Hagen2009; @Mehr2021; @Mehr2017]. These constraints on cultural-evolutionary processes result in musical behaviours that show elements of cultural specificity while still remaining grounded in general biological tendencies [@Richerson2008; @Sperber2004]. The resulting regularities enable listeners to reliably infer the behavioural contexts of unfamiliar foreign music [@Mehr2018a; @Mehr2019], even young children, who have less musical experience relative to adults [@Hilton2022b].

Notably, while prior experiments demonstrate that people can infer the behavioural contexts of songs from different cultures using their acoustic features, these studies frequently have sampling limitations. For instance, some studies rely primarily on English-speaking Western participants [@Trehub1993], and those that have reached participants around the world still rely on English speakers who have access to the Internet [@Mehr2018a; @Mehr2019; @Hilton2022a]; n.b., this important problem affects many areas of the cognitive sciences [@Blasi2022]. Thus, although the stimuli participants in these studies listened to were cross-culturally representative, it is unclear how much of the accuracy of listener inferences is accounted for by universal form-function links in musical behaviour, and how much is a product of (Western) enculturation, education, and exposure to world music through globalised media.

Here, we test the prediction that the behavioural contexts of songs are mutually intelligible to listeners across cultures. We study a large and diverse sample of listeners recruited worldwide in many languages, from both industrialised societies and smaller-scale societies. We use *smaller-scale* to refer to (i) societies in which individuals interact in a "small" world (i.e., 10-100 other individuals but not more), most interactions are face-to-face, and there is a high degree of interdependence); and (ii) societies less affected by states, markets, globalization, and/or world religions.

We predicted that listeners in both industrialised and smaller-scale societies would correctly infer the behavioural contexts of three types of unfamiliar songs (dance, lullaby, healing), reflecting a sensitivity to acoustic cues shared in these contexts across cultures (the preregistration is at [https://osf.io/msvwz](https://osf.io/msvwz)). In exploratory analyses, we asked whether culturally learned cues would give listeners an advantage when inferring the behavioural contexts of songs that are more closely related to their own culture, in line with other domains, such as the perception of emotion in vocalizations [@Laukka2021; @Elfenbein2002].

@@

Note that in the cases of countries with multiple official languages, we were not always successful in our goal of only recruiting native speakers of non-English languages, due to recruitment difficulties. As a result, some participants in some countries were split across native language groupings. For example, the South African sample included native speakers of both Zulu and English (contrary to our plan to include only native speakers of Zulu), whereas the Kenyan sample included only native speakers of Swahili (as planned). Further details on deviations from the preregistered recruitment plan are in SI Text 1.1.

We aimed to maximise data quality with eight planned exclusion criteria: we excluded participants who (i) performed poorly on a commonly used headphone detection task [@Woods2017]; (ii) reported difficulties hearing the audio on at least 4 of 24 trials (e.g., because of poor connectivity); (iii) had an IP address that did not geolocate to the same country they reported as their location; (iv) failed a simple attention check; (v) completed the survey more rapidly than would be possible; (vi) reported not wearing headphones; (vii) reported being in a noisy environment; or (viii) reported not being careful in completing the study. After exclusions, the sample included n = `r info$web$n %>% f` native speakers of `r info$web$lang %>% f` languages, located in `r info$web$country %>% f` countries.

Qualtrics Panels compensated each participant directly in the local currency, with rates varying across countries as a function of local payment standards for survey participation. All participants provided informed consent, under a protocol approved by the Harvard University Committee on the Use of Human Subjects (Ethics ID: IRB16-1080).

\renewcommand{\arraystretch}{1.3}
```{r table1-manual, results='asis'}
## temporary table 1 code -- to update to procedural later
manual_table1 <- read.csv(here("viz","table1.csv"))
opts <- options(knitr.kable.NA="")

kable(manual_table1,
      format = "latex",
      booktabs = TRUE,
      escape = FALSE,
      linesep = "",
      col.names = c("Language family","Language","Total $n$","Subregion","Country","Country-wise $n$"),
      longtable = TRUE,
) %>% 
  # partial horizontal lines
  row_spec(c(1,8,10,11,14,15,16,18,23,25,35,37,38,39,41,42,43,47,48), extra_latex_after = "\\cline{2-6}") %>%
  row_spec(c(3,6,21,24,26:30,33,34,36,40,44,45,46,54), extra_latex_after = "\\cline{4-6}") %>%
  # full horizontal lines
  row_spec(c(7,9,12,49,50,51,53,55), hline_after = TRUE) %>%
  kable_styling(font_size = 8,
                full_width = FALSE) %>%
  footnote(general = "Linguistic and geographic information about the participants in the industrialised societies. Language information refers to the native language of the participant; languages were classified via Glottolog. Geographic information refers to the country of residence of the participant; world subregions were classified via the Human Relations Area Files.",
           general_title = "Table 1.",
           footnote_as_chunk = TRUE,
           title_format = "bold",
           escape = FALSE,
           threeparttable = TRUE)

@@

```{r supplementary-table-1a, results='asis', eval=FALSE}
## code for generating table 1 data, not currently in use and for implementing with procedural table 1 later
tmp <- webraw %>% 
  select(indx, natlang_name, natlang_fam, natcountry, listener_subregion) %>% 
  distinct() %>% 
  mutate(natlang_fam = recode(natlang_fam,
                indo1319 = "Indo-European",
                afro1255 = "Afro-Asiatic",
                turk1311 = "Turkic",
                atla1278 = "Atlantic-Congo",
                aust1307 = "Austronesian",
                ural1272 = "Uralic",
                japo1237 = "Japonic",
                kore1284 = "Koreanic",
                sino1245 = "Sino-Tibetan",
                aust1305 = "Austroasiatic"
  )) %>% 
  group_by(natlang_name) %>% 
  mutate(total_n = n()) %>% 
  group_by(natlang_name, natcountry) %>% 
  mutate(`Country-wise n` = n()) %>% 
  select(
    `Language Family` = natlang_fam,
     Language = natlang_name,
    `Total n` = total_n,
    Subregion = listener_subregion,
    Country = natcountry,
    `Country-wise n`
  ) %>% distinct() %>% 
  arrange(`Language Family`, Language, Subregion, Country) %>% 
  write.csv(here("viz","table1.csv"))

## temporary code here to manually prettify this table -- to update later
manual_table1 <- read.csv(here("viz","table1.csv"))
  kable(manual_table1,
        format = "latex",
        booktabs = TRUE,
        escape = FALSE,
        linesep = "\\addlinespace",
        longtable = TRUE,
        align=rep('l', 6))%>%
  kable_styling(font_size = 7.5,
                full_width = FALSE) %>%
  column_spec(1, width = "1in") %>%
  column_spec(2, width = "1.2in") %>%
  column_spec(3, width = "0.5in") %>% 
  kableExtra::collapse_rows(columns = c(1,2,3,4),
                            longtable_clean_cut= T,
                            target = 2,
                            custom_latex_hline=5) %>% 
  footnote(general = "Linguistic and geographic information about the participants in the web-experiment. The 'Language' column denotes the native language spoken by the participant (and the language they completed the experiment in); the 'Total n' column denotes the number of participants recruited for that language; the 'Language Family' column denotes the language family this particular language is part of, as per the Glottolog system [@Hammarstrom2019]. Glottolog is a comprehensive catalogue of the world's languages and their genealogy, and can be accessed at [glottolog.org](https://glottolog.org/). Within each language, participants were recruited from multiple countries, as noted in the 'Country' column. For the cultural proximity analyses, participants were then grouped into geographic subregions based on their reported location, as per the Human Relations Area Files World Sub-region typology. The country-wise n indicates the number of participants per language x country.",
           general_title = "Table 1.",
           footnote_as_chunk = TRUE,
           title_format = "bold",
           escape = FALSE,
           threeparttable = TRUE)

@@

We recruited adult participants from the Nyangatom in Ethiopia (*n* = 35), the Mentawai in Indonesia (*n* = 30), and the Tannese Ni-Vanuatu in Vanuatu (*n* = 56), via word-of-mouth sampling. The approximate locations of each of these smaller-scale societies are visualised in Figure 1 and summary information about each is in Table 2. The societies were chosen for their reduced exposure to music from other cultural traditions. At the time of data collection (2017 to 2019), all three societies had somewhat limited access to TV, radio and the Internet and could not be assumed to have had significant exposure to these communication channels.^[At the time of data collection, the three societies had somewhat limited access to TV, radio and the Internet and they varied in their familiarity with these technologies. The Nyangatom communities had little exposure to these technologies when the experiment was conducted, although exposure has since expanded considerably. The Ni-Vanuatu communities were exposed to Christian music in church, as well as reggae and other foreign music through battery powered radios and, over the last five years, increasing access to the Internet via cell phones. Nonetheless, traditional Kastom music is still widely performed in local religious and civil ceremonies and is an important part of Ni-Vanuatu culture and identity. The Mentawai communities studied encountered non-Mentawai music, particularly Indonesian and Bollywood music, through both radios and memory sticks purchased in the port-town, although both cell phone and radio ownership were rare.] In each society, indigenous music continues to be widespread and central to cultural identity.

In the cases of 5 participants, an experimenter expressed concern as to whether the participant understood the task; these participants were excluded without the experimenter being aware of the songs heard. As in the industrialised cohort, participants were compensated directly in the local currency, with rates determined by the Principal Investigator at each site and in keeping with norms across other research projects conducted in the area. Ethics approval was granted by the Pennsylvania State University Office for Research Protections (Ethics ID: STUDY00012265) for data collection in Ethiopia; the Institute for Advanced Study in Toulouse (Ethics ID: 2017-09-001) for data collection in Indonesia; and the University of Auckland Human Participants Ethics Committee (Ethics ID: 021538) for data collection in Vanuatu.

```{r fig1, eval=FALSE, fig.cap="\\textbf{Figure 1 | Geographic distribution of participants.} We recruited participants in industrialised societies and in three smaller-scale societies. The grey dots indicate the approximate locations of the participants in industrialised societies, fig.width=6, as measured via IP geolocation. The yellow dots indicate the approximate locations of the three smaller-scale societies (from left to right, the Nyangatom, Mentawai Islanders, and Tannese Ni-Vanuatu).", fig.height=3, include=FALSE}

>>>>

---
bibliography: themusiclab.bib
csl: pnas.csl
header-includes:
- \usepackage[left]{lineno}
- \usepackage{ragged2e}
- \usepackage{caption}
- \usepackage{longtable}
- \usepackage[labelformat = empty]{caption}
- \usepackage{afterpage}
- \usepackage{mdframed}
- \usepackage{fontenc}
- \usepackage{soul}
- \usepackage{xcolor}
- \usepackage[symbol]{footmisc}
- \definecolor{bleu}{HTML}{2200cc}
- \renewcommand{\thefootnote}{\fnsymbol{footnote}}
- \usepackage{rotating}
notes-after-punctuation: no
urlcolor: bleu
linkcolor: bleu
link-citations: yes
output:
  pdf_document:
    number_sections: yes
  word_document:  
  html_document:
    df_print: paged
---

@@

```
```{r libraries}

# note, make sure {plyr} is installed: install.packages("plyr")

library(pacman)
p_load(
  here,
  nlme,
  lme4,
  multcomp,
  RColorBrewer,
  lmerTest,
  gt,
  scales,
  ggpubr,
  geodist,
  cowplot,
  patchwork,
  ggsignif,
  broom,
  broom.mixed,
  glmnet,
  ggeffects,
  ggtext,
  ggnewscale,
  glue,
  rgeos,
  rnaturalearth,
  rnaturalearthdata,
  sf,
  tidyverse,
  kableExtra,
  factoextra,
  tidymodels
)

load(here("results", "analysis.RData"))
load(here("results", "lasso_analyses.RData"))

@@

\raggedright
\LARGE
\textbf{Universal interpretations of vocal music}

@@

\justifying
\normalsize
Lidya Yurdum^1,2,$\ast$^, Manvir Singh^3^, Luke Glowacki^4^, Thomas Vardy^5^, Quentin D. Atkinson^5^, Courtney B. Hilton^5^, Disa Sauter^2^, Max M. Krasnow^6^, \& Samuel A. Mehr^1,5,$\ast$^

\small
^1^Child Study Center, Yale University, New Haven, CT 06520, USA.  
^2^Department of Psychology, University of Amsterdam, Amsterdam 1018WT, Netherlands.  
^3^Institute for Advanced Study in Toulouse, 31080 Toulouse Cedex 6, France.  
^4^Department of Anthropology, Boston University, Boston, MA 02215, USA.  
^5^School of Psychology, University of Auckland, Auckland 1010, New Zealand.  
^6^Division of Continuing Education, Harvard University, Cambridge, MA 02138, USA.

\*Corresponding authors. E-mail: [lidya.yurdum\@yale.edu](mailto:lidya.yurdum@yale.edu){.email}, [sam\@auckland.ac.nz](mailto:sam@auckland.ac.nz){.email}

@@

```{=tex}
\normalsize
\begin{mdframed}[backgroundcolor=gray!20]
Despite the variability of music across cultures, some types of human songs share acoustic characteristics. For example, dance songs tend to be loud and rhythmic and lullabies tend to be quiet and melodious. Human perceptual sensitivity to the behavioural contexts of songs, on the basis of these musical features, raises the possibility that basic properties of music are mutually intelligible, independent of linguistic or cultural content. Whether these effects reflect universal interpretations of vocal music, however, is unclear, because prior studies focus almost exclusively on English-speaking participants, a group that is not representative of humans writ large. Here we report shared intuitions concerning the behavioural contexts of unfamiliar songs produced in unfamiliar languages, in participants living in Internet-connected industrialised societies ($n = 5,516$ native speakers of 28 languages) or smaller-scale societies with limited access to global media ($n = 116$ native speakers of 3 non-English languages). Participants listened to songs randomly selected from a representative sample of human vocal music, originally used in four behavioural contexts, and rated the degree to which they believed the song was used for each context. Listeners in both industrialised and smaller-scale societies reliably inferred the contexts of dance songs, lullabies, and healing songs, but not love songs. Within and across the cohorts, inferences were mutually consistent. Further, increased linguistic or geographical proximity between listeners and singers only minimally increased the accuracy of the inferences. These results demonstrate that the behavioural contexts of three common forms of music are mutually intelligible across cultures and imply that musical diversity, shaped by cultural evolution, is nonetheless grounded in some universal perceptual phenomena.

@@

Like many other animals, humans use vocalisations to convey their intentions and affective states [@Morton1977; @Pisanski2022]. Such vocalisations would be meaningless in a world where members of one's own — or other — species could not interpret these signals in a useful way. Indeed, many animal and human vocalisations are not arbitrary but instead display systematic relationships between their acoustic form and their behavioural function [@Endler1993; @Fitch2002; @Pisanski2022]. For instance, the human scream is unlikely to have evolved arbitrarily as a means of communicating distress and urgency: rather, a scream involves extreme high frequencies [@Pisanski2020] and acoustic roughness [@Arnal2015] that set it apart from regular verbal communication, and make it appropriate for the behavioural function of grabbing attention.

@@

Systematic form-function relationships also occur in more complex vocalisations. Vocal music (hereafter, song) is a human universal characterised by rich variability within and across cultures [@Mehr2019; @Nettl1977; @Lomax1968]. Some of the behavioural contexts in which songs are used, however, are conspicuously similar around the globe, such as singing to soothe fussy infants or singing to coordinate dancing [@Trehub1993; @Trehub1993a; @Mehr2018a; @Mehr2019; @Hilton2022a; @Hilton2022b; @Singh2023]. Songs used for specific functions in specific behavioural contexts tend to have objective acoustic correlates; that is, they tend to display stereotyped musical features associated with their specific behavioural context. For example, dance songs tend to share clearly accented and predictable beat structures.

As with other types of vocalisations, form-function patterns in human song may originate from our evolved psychology, perceptual biases, or unique social environment [@Hagen2003; @Hagen2009; @Mehr2021; @Mehr2017]. These constraints on cultural-evolutionary processes result in musical behaviours that show elements of cultural specificity while still remaining grounded in general biological tendencies [@Richerson2008; @Sperber2004]. The resulting regularities enable listeners to reliably infer the behavioural contexts of unfamiliar foreign music [@Mehr2018a; @Mehr2019], including young children, who have less musical experience relative to adults [@Hilton2022b].

While prior experiments have shown that people can infer the behavioural contexts of songs from different cultures using only acoustic features of the songs, these studies frequently have sampling limitations. For instance, some studies rely primarily on English-speaking Western participants [@Trehub1993], and those that have reached participants around the world still rely on English speakers who have access to the Internet [@Mehr2018a; @Mehr2019; @Hilton2022a] — an important problem affecting many areas of the cognitive sciences [@Blasi2022]. Thus, although the stimuli participants in these studies listened to were cross-culturally representative, it is unclear how much of the accuracy of listener inferences is accounted for by universal form-function links in musical behaviour, and how much is a product of (Western) enculturation, education, and exposure to world music through globalised media.

Here, we test the prediction that the behavioural contexts of songs are mutually intelligible to listeners across cultures. We study a large and diverse sample of listeners recruited worldwide in many languages, from both industrialised societies and smaller-scale societies. We use *smaller-scale* to refer to (i) societies in which individuals interact in a "small" world (i.e., 10-100 other individuals but not more), most interactions are face-to-face, and there is a high degree of interdependence; and (ii) societies less affected by states, markets, globalization, and/or world religions.

We predicted that listeners in both industrialised and smaller-scale societies would correctly infer the behavioural contexts of three types of unfamiliar songs (dance, lullaby, healing), reflecting a sensitivity to acoustic and musical cues shared in these contexts across cultures (the preregistration is at [https://osf.io/msvwz](https://osf.io/msvwz)). In exploratory analyses, we asked whether culturally learned cues would give listeners an advantage when inferring the behavioural contexts of songs that are more closely related to their own culture, in line with other domains, such as the perception of emotion in vocalizations [@Laukka2021; @Elfenbein2002].

@@

In the cases of countries with multiple official languages, we were not always successful in our goal of only recruiting native speakers of non-English languages, due to recruitment difficulties. As a result, some participants in some countries were split across native language groupings. For example, the South African sample included native speakers of both Zulu and English (contrary to our plan to include only native speakers of Zulu), whereas the Kenyan sample included only native speakers of Swahili (as planned). Further details on deviations from the preregistered recruitment plan are in SI Text 1.1.

We aimed to maximise data quality with eight planned exclusion criteria: we excluded participants who (i) performed poorly on a headphone detection task [@Woods2017]; (ii) reported difficulties hearing the audio on at least 4 of 24 trials (e.g., because of poor connectivity); (iii) had an IP address that did not geolocate to the same country they reported as their location; (iv) failed a simple attention check; (v) completed the survey more rapidly than should be possible; (vi) reported not wearing headphones; (vii) reported being in a noisy environment; or (viii) reported not being careful in completing the study. After exclusions, the sample included `r info$web$n %>% f` native speakers of `r info$web$lang %>% f` languages, located in `r info$web$country %>% f` countries.

Qualtrics Panels compensated each participant directly in the local currency, with rates varying across countries as a function of local payment standards for survey participation. All participants provided informed consent. The study protocol was approved by the Harvard University Committee on the Use of Human Subjects (protocol IRB16-1080).

\renewcommand{\arraystretch}{1.3}

```{r table1-manual, results='asis'}
## temporary table 1 code -- to update to procedural later
manual_table1 <- read.csv(here("viz","table1.csv"))
opts <- options(knitr.kable.NA="")

kable(manual_table1,
      format = "latex",
      booktabs = TRUE,
      escape = FALSE,
      linesep = "",
      col.names = c("Language family","Language","Total $n$","Subregion","Country","Country-wise $n$"),
      longtable = TRUE,
) %>% 
  # partial horizontal lines
  row_spec(c(1,8,10,11,14,15,16,18,23,25,35,37,38,39,41,42,43,47,48), extra_latex_after = "\\cline{2-6}") %>%
  row_spec(c(3,6,21,24,26:30,33,34,36,40,44,45,46,54), extra_latex_after = "\\cline{4-6}") %>%
  # full horizontal lines
  row_spec(c(7,9,12,49,50,51,53,55), hline_after = TRUE) %>%
  kable_styling(font_size = 8,
                full_width = FALSE) %>%
  footnote(general = "Linguistic and geographic information about the participants in the web-experiment. The ``Language'' column denotes the native language spoken by the participant (and the language they completed the experiment in); the ``Total n'' column denotes the number of participants recruited for that language; the ``Language Family'' column denotes the language family this particular language is part of, as per the Glottolog system (40). Glottolog is a comprehensive catalogue of the world's languages and their genealogy, and can be accessed at https://glottolog.org. Within each language, participants were recruited from multiple countries, as noted in the ``Country'' column. For the cultural proximity analyses, participants were then grouped into geographic subregions based on their reported location, as per the Human Relations Area Files World Sub-region typology. The ``Country-wise n'' column indicates the number of participants per language in each country.",
           general_title = "Table 1.",
           footnote_as_chunk = TRUE,
           title_format = "bold",
           escape = FALSE,
           threeparttable = TRUE)

@@

```{r supplementary-table-1a, eval=FALSE, include=FALSE, results='asis'}
## code for generating table 1 data, which we then edited manually to format
tmp <- webraw %>% 
  select(indx, natlang_name, natlang_fam, natcountry, listener_subregion) %>% 
  distinct() %>% 
  mutate(natlang_fam = recode(natlang_fam,
                indo1319 = "Indo-European",
                afro1255 = "Afro-Asiatic",
                turk1311 = "Turkic",
                atla1278 = "Atlantic-Congo",
                aust1307 = "Austronesian",
                ural1272 = "Uralic",
                japo1237 = "Japonic",
                kore1284 = "Koreanic",
                sino1245 = "Sino-Tibetan",
                aust1305 = "Austroasiatic"
  )) %>% 
  group_by(natlang_name) %>% 
  mutate(total_n = n()) %>% 
  group_by(natlang_name, natcountry) %>% 
  mutate(`Country-wise n` = n()) %>% 
  select(
    `Language Family` = natlang_fam,
     Language = natlang_name,
    `Total n` = total_n,
    Subregion = listener_subregion,
    Country = natcountry,
    `Country-wise n`
  ) %>% distinct() %>% 
  arrange(`Language Family`, Language, Subregion, Country) %>% 
  write.csv(here("viz","table1_notFormatted.csv"))

@@

We recruited adult participants from the Nyangatom in Ethiopia (*n* = 35), the Mentawai in Indonesia (*n* = 30), and the Tannese Ni-Vanuatu in Vanuatu (*n* = 56), via word-of-mouth sampling. The approximate locations of each of these smaller-scale societies are visualised in Figure 1 and summary information about each is in Table 2. The societies were chosen for their reduced exposure to music from other cultural traditions. At the time of data collection (2017 to 2019), all three societies had somewhat limited access to TV, radio and the Internet and could not be assumed to have had significant exposure to these communication channels.^[The Nyangatom communities had little exposure to TV, radio, and the Internet when the experiment was conducted, although exposure has since expanded considerably. The Ni-Vanuatu communities were exposed to Christian music in church, as well as reggae and other foreign music through battery powered radios and, over the last five years, increasing access to the Internet via cell phones. Nonetheless, traditional Kastom music is still widely performed in local religious and civil ceremonies and is an important part of Ni-Vanuatu culture and identity. The Mentawai communities studied encountered non-Mentawai music, particularly Indonesian and Bollywood music, through both radios and memory sticks purchased in the port-town, although both cell phone and radio ownership were rare.] In each society, indigenous music continues to be widespread and central to cultural identity.

In the cases of 5 participants, an experimenter expressed concern as to whether the participant understood the task; these participants were excluded without the experimenter being aware of the songs heard. As in the industrialised cohort, participants were compensated directly in the local currency, with rates determined by the Principal Investigator at each site and in keeping with norms across other research projects conducted in the area. Ethics approval was granted by the Pennsylvania State University Office for Research Protections (protocol STUDY00012265) for data collection in Ethiopia; the Institute for Advanced Study in Toulouse (protocol 2017-09-001) for data collection in Indonesia; and the University of Auckland Human Participants Ethics Committee (protocol 021538) for data collection in Vanuatu.

```{r fig1, eval=FALSE, fig.cap="\\textbf{Figure 1 | Geographic distribution of participants.} We recruited participants in industrialised societies and in three smaller-scale societies. The black dots indicate the approximate locations of the participants in industrialised societies, fig.width=6, as measured via IP geolocation. The yellow dots indicate the approximate locations of the three smaller-scale societies (from left to right, the Nyangatom, Mentawai Islanders, and Tannese Ni-Vanuatu).", fig.height=3, include=FALSE}
